---
title: "GLM"
author: "DAS Group 07"
date: "19/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE,echo=TRUE}
library(tidyverse)
library(kableExtra)
library(knitr)
library(skimr)
library(ggcorrplot)
library(gridExtra)
library(countrycode)
library(car)
library(moderndive)
library(jtools)
```

```{r data, echo= FALSE, eval = TRUE}
# import and process data
coffee <- read.csv('dataset7.csv')
```

```{r process, echo= FALSE, warning=FALSE}
coffee <- as_tibble(coffee)
coffee <- coffee %>% 
  rename(altitude = altitude_mean_meters) %>% 
  rename(defects = category_two_defects) %>% 
  rename(country = country_of_origin)
# removing the two data errors for altitude
coffee$altitude <- ifelse(coffee$altitude > 8000, NA, coffee$altitude)
#removing na obs
coffee <- coffee %>% na.omit()
coffee <- as.data.frame(coffee)
```

```{r add variable, echo= FALSE, warning=FALSE}
coffee$continent = countrycode(sourcevar = coffee[,"country"],
                               origin="country.name",
                               destination = "continent")
coffee[c(374,408,764,787),"continent"] <- 'Americas'
coffee <- as_tibble(coffee)
```

# Formal Data Analysis 

In formal data analysis, we use three link functions to fit generalized linear models, then use step-wise regression to select reasonable explanatory variables based on AIC values, and finally use chi-square values to judge the fitness of the model.

```{r model selection, echo= FALSE, warning=FALSE}
coffee1 <- as.data.frame(coffee)

coffee1[which(coffee1$Qualityclass=="Poor"),]$Qualityclass=0
coffee1[which(coffee1$Qualityclass=="Good"),]$Qualityclass=1

coffee1$Qualityclass = as.factor(coffee1$Qualityclass)
```


```{r logit, echo= FALSE, warning=FALSE}
logit_model = glm(Qualityclass~aroma + flavor + acidity + defects + altitude + harvested + continent, data = coffee1, family = binomial(link = "logit"))
both_logit = step(logit_model,direction="both")
summ(both_logit)
both_logit$anova %>%
  select(Step, AIC) %>%
  kable() %>%
  kable_styling(font_size = 10, latex_options = "hold_position")

## formula = Qualityclass ~ aroma + flavor + acidity + altitude + harvested
## AIC: 533.47, BIC = 562.23

## Fitness of model
summary(both_logit)$null.deviance - summary(both_logit)$deviance > qchisq(0.95,891-886)
## TRUE. we can reject the null hypothesis, and the terms are all significant
```

```{r probit, echo= FALSE, warning=FALSE}

# Probit link
probit_model = glm(Qualityclass~aroma + flavor + acidity + defects + altitude + harvested + continent, data = coffee1, family = binomial(link = "probit"))
both_probit = step(probit_model,direction = "both")
summ(both_probit)
both_probit$anova %>%
  select(Step, AIC) %>%
  kable() %>%
  kable_styling(font_size = 10, latex_options = "hold_position")

## formula = Qualityclass ~ aroma + flavor + acidity + altitude + harvested
## AIC = 554.03, BIC = 582.79
## Fittness of model
summary(both_probit)$null.deviance - summary(both_probit)$deviance > qchisq(0.95,891-886)
## TRUE. we can reject the null hypothesis, and the terms are all significant

```


```{r cloglog, echo= FALSE, warning=FALSE}

clog_model = glm(Qualityclass~aroma + flavor + acidity + defects + altitude + harvested + continent, data = coffee1, family = binomial(link = "cloglog"))
both_clog = step(clog_model,direction = "both")
summ(both_clog)
both_clog$anova %>%
  select(Step, AIC) %>%
  kable() %>%
  kable_styling(font_size = 10, latex_options = "hold_position")

## formula = Qualityclass ~ aroma + flavor + acidity + harvested
## AIC = 636.96, BIC = 660.93
## Fittness of model
summary(both_clog)$null.deviance - summary(both_clog)$deviance > qchisq(0.95,891-887)
## TRUE. we can reject the null hypothesis, and the terms are all significant

```

According to the stepwise regression results, the GLM explanatory variables of complementary log-log link are *aroma*, *flavor*, *acidity* and *harvested*, and the GLM explanatory variables of logit link and probit link are *aroma*, *flavor*, *acidity*, *altitude* and *harvested*.

The Pearson chi-squared statistics of three models are all greater than the 95th percentile of the $\chi^{2}(4)$ distribution. Therefore the models fit the data well and We need to choose the appropriate link function by comparing the information criteria.

# GLM Model Selection

The AIC and BIC values corresponding to each link function are as follows:

```{r aicbic, echo= FALSE, warning=FALSE}
summlogit = summ(both_logit)
summprobit = summ(both_probit)
summclog = summ(both_clog)
aic1 <- round(summlogit$model$aic, 3)
aic2 <- round(summprobit$model$aic, 3)
aic3 <- round(summclog$model$aic, 3)
```
 
 Link                      | Link Function   | AIC      | BIC   
 :-------------------------|:--------------- |:--------:|:-----:
 Logit link                | $g\left(p_{i}\right)=\log \left(\frac{p_{i}}{1-p_{i}}\right)$             | 533.47   | 562.23    
 Probit link               | $g\left(p_{i}\right)=\Phi^{-1}\left(p_{i}\right)=\beta_{0}+\beta_{1} x_{i}$             | 554.03   | 582.79   
 Complementary log-log link| $g\left(p_{i}\right)=\log \left[-\log \left(1-p_{i}\right)\right]=\beta_{0}+\beta_{1} x_{i}$             | 636.96    | 660.93  

Based on the AIC and BIC values in the table above, the model using logit link fits best in three. So we finally choose the logit link function in GLM.

The GLM regression model of logit link is as follows:

$$
Y \sim B(m_i, p{(\text {Qualityclass = Good})}_i),
$$

$$
g\left(p{(\text {Qualityclass = Good})}_{i}\right)=\log \left(\frac{p{(\text {Qualityclass = Good})}_{i}}{1-p{(\text {Qualityclass = Good})}_{i}}\right),
$$

```{r logitformula, echo= FALSE, warning=FALSE}
logitsele = summary(both_logit)
Coefs <- round(coef(logitsele), 4)
```

$$
\log \left(\frac{p{(\text {Qualityclass = Good})}}{1-p{(\text {Qualityclass = Good})}}\right) = `r Coefs[1]` + `r Coefs[2]` \cdot aroma + `r Coefs[3]` \cdot flavor + `r Coefs[4]` \cdot acidity + `r Coefs[5]` \cdot altitude + `r Coefs[6]` \cdot harvested.
$$



Considering that the correlation coefficient of aroma, flavor and acidity in EDA is relatively large, we calculated the VIF value of the variables in the regression. 

```{r VIF, echo= FALSE, warning=FALSE}
vif(both_logit) %>%
  kable(caption = '\\label{tab:VIF} VIF of Variables', digits = 2)%>%
  kable_styling(latex_options = "hold_position")
```

The results show that the VIF values are all small, excluding multicollinearity.


