---
title: "Group_7_EDF_Fork"
author: "DAS Group 07"
date: "15/03/2022"
output:
  slidy_presentation: default
  beamer_presentation: default
  powerpoint_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(kableExtra)
library(skimr)
library(ggcorrplot)
library(gridExtra)
library(countrycode)
library(sjPlot)
```

```{r data and packages, include=FALSE}
coffee <- read.csv('dataset7.csv')
coffee <- as_tibble(coffee)
coffee <- coffee %>% 
  rename(altitude = altitude_mean_meters) %>% 
  rename(defects = category_two_defects) %>% 
  rename(country = country_of_origin)
glimpse(coffee)
summary(coffee)
```

Initially, summary statistics reveal a couple of erroneous altitude datapoints. (The highest two values are many times the altitude of the peak of Mount Everest!)

```{r}
# removing the two data errors for altitude
coffee$altitude <- ifelse(coffee$altitude > 8000, NA, coffee$altitude)
#removing na obs
coffee <- coffee %>% na.omit()
```


Table of summary statistics for continuous variables:

```{r cont_summaries, warning = FALSE}
cont_summ <- coffee %>% 
  select(aroma, flavor, acidity, altitude, defects) %>% 
  summarise_each(funs(mean = round(mean(., na.rm = TRUE), 2),
                      sd = round(sd(., na.rm = TRUE), 2),
                      min = min(., na.rm = TRUE),
                      q25 = quantile(., 0.25, na.rm = TRUE), 
                      median = median(., na.rm = TRUE), 
                      q75 = quantile(., 0.75, na.rm = TRUE), 
                      max = max(., na.rm = TRUE))) %>% 
  gather(stat, val) %>%
  separate(stat, into = c("var", "stat"), sep = "_") %>%
  spread(stat, val) %>%
  select(var, mean, sd, min, q25, median, q75, max) %>%
  slice(3, 5, 1, 2, 4) %>% 
  kable(caption = 'Summary statistics for coffee grades (1-10), mean altitude (metres) and defects (integer).')

cont_summ
```

Summary statistics reveal a couple of erroneous altitude datapoints. (The highest two values are many times the altitude of the peak of Mount Everest!)


A correlation plot of the numeric variables was generated:

```{r}
correlations <- cor(coffee[,2:6], use="pairwise.complete.obs") # ignore missing values
ggcorrplot(correlations, 
           hc.order = TRUE, type = "lower", lab = TRUE)
```

The three coffee scoring variables (aroma, flavour, acidity) seem strongly positive correlated, so we may not expect to see all in a final model.

Boxplots of the numerical variables were produced, split by quality classification:

```{r boxplots, warning = FALSE}
box1 <- coffee %>%
  ggplot(aes(x = Qualityclass, y = aroma)) +
  geom_boxplot(aes(fill = Qualityclass), show.legend = FALSE) +
  labs(x = "Quality Class", y = "Aroma grade") +
  theme(axis.title = element_text(size = 9))

box2 <- coffee %>%
  ggplot(aes(x = Qualityclass, y = flavor)) +
  geom_boxplot(aes(fill = Qualityclass), show.legend = FALSE) +
  labs(x = "Quality Class", y = "Flavour grade") +
  theme(axis.title = element_text(size = 9))

box3 <- coffee %>%
  ggplot(aes(x = Qualityclass, y = acidity)) +
  geom_boxplot(aes(fill = Qualityclass), show.legend = FALSE) +
  labs(x = "Quality Class", y = "Acidity grade") +
  theme(axis.title = element_text(size = 9))

box4 <- coffee %>%
  ggplot(aes(x = Qualityclass, y = altitude)) +
  geom_boxplot(aes(fill = Qualityclass), show.legend = FALSE) +
  labs(x = "Quality Class", y = "Altitude (m)") +
  theme(axis.title = element_text(size = 9))

box5 <- coffee %>%
  ggplot(aes(x = Qualityclass, y = defects)) +
  geom_boxplot(aes(fill = Qualityclass), show.legend = FALSE) +
  labs(x = "Quality Class", y = "# of defects") +
  theme(axis.title = element_text(size = 9))

box6 <- coffee %>%
  ggplot(aes(x = Qualityclass, y = harvested)) +
  geom_boxplot(aes(fill = Qualityclass), show.legend = FALSE) +
  labs(x = "Quality Class", y = "Year") +
  theme(axis.title = element_text(size = 9))

lay <- cbind(c(1,1,2,2,3,3),
             c(4,4,5,5,6,6))
grid.arrange(box1, box2, box3, box5, box4, box6, layout_matrix = lay)
```

```{r conti var, warnings = FALSE}

coffee <- as.data.frame(coffee)
coffee$continent = countrycode(sourcevar = coffee[,"country"],
                         origin="country.name",
                         destination = "continent")
# United States (Puerto Rico) were not matched unambiguously
coffee[c(374,408,764,787),"continent"] <- 'Americas'
coffee <- as_tibble(coffee)

```

Bar plots of quality by country and year.

```{r bar charts, warnings = FALSE}

countries <- table(coffee$Qualityclass, coffee$country)
countries <- as.data.frame(countries)
colnames(countries)<- c('quality','country','count')

bar1 <- countries %>% 
  ggplot(aes(x = country, y = count)) +
  geom_col(aes(fill = quality)) +
  coord_flip()

bar1

continent <- table(coffee$Qualityclass, coffee$continent)
continent <- as.data.frame(continent)
colnames(continent)<- c('quality','continent','count')

bar2 <- continent %>% 
  ggplot(aes(x = count, y = continent)) +
  geom_col(aes(fill = quality)) +
  coord_flip()

bar2

#Easy to compare the quality of coffee beans within the same continent
plot_xtab(coffee$continent,coffee$Qualityclass, show.values = FALSE,
          show.total = FALSE,
          legend.title = "Quality")

```

